---
title: "Assessing the Impact of Affordable Housing Development"
output: html_notebook
---
 


```{r}
#install.packages("sf")
#install.packages("leaflet")
#install.packages("lme4")
#install.packages("AER")
#install.packages("MASS")
library(sf)
library(tidyverse) 
library(lme4)
library(MASS)
library(AER)
library(lme4)
library(leaflet)
library(lubridate)

```

Let's load in our zipcode dataset. This is a geojson file, which was downloaded from <https://data.nashville.gov/Metro-Government/Zip-Codes/72um-tmbe>. Geosjson is a standard, based on JSON, for representing geographic features.

```{r}
CensusTracts <- st_read('../data/DC')
```

Notice that the ouput provides some information about the data we just read in. It contains multipolygons and is using WGS 84 as the CRS, or coordinate reference system, which defines how the points on the globe are mapped to the 2-dimensional projection we are working with.

The actual object we read in is a simple feature collection, which means that it acts like a data frame, but it also contains additional geographic information.

```{r}
CensusTracts
```

We can create plots out of our geospatial data using ggplot.

```{r}
CensusTracts |> 
  ggplot() +
  geom_sf() 
```

 

We might also want to work with data in the form a csv file.

```{r}
Sales <- read_csv("../data/filtered_sales.csv")
Sales
```

```{r}
LIHTC <- read_csv("../data/LIHTC.csv")
LIHTC
```

```{r}

AffordableHousing <- LIHTC |> filter(HUD_ID=='TNA20130015')
AffordableHousing 


```


```{r}
PropertyDetails <- read_csv("../data/property_details.csv")
PropertyDetails
```

```{r}
Price_n_Location <- Sales |> 
                      left_join(PropertyDetails, by = join_by(apn)) |> 
                      rename(price=amount)
Price_n_Location 
```


 

```{r}

Price_n_Location <- Price_n_Location |>  
                      mutate(across('centroid', str_replace, '\\(', '')) |> 
                      mutate(across('centroid', str_replace, '\\)', '')) |> 
                      separate(`centroid`, sep = ",", into = c("LONGITUDE", "LATITUDE"))   |> 
                      mutate(across(LONGITUDE:LATITUDE, as.numeric)) |> 
                      relocate(LONGITUDE, .after = LATITUDE)
Price_n_Location
```








Now, we can plot these burglary incidents. Note that I'll start with the census tract plot and then add in the burglary incidents by using geom_point.

```{r}
CensusTracts |>
  ggplot() +
  geom_sf() +
  geom_point(
    data = Price_n_Location |> drop_na(LATITUDE),
    aes(x = LONGITUDE, y = LATITUDE),
    size = 0.1
    )
```
 

 

```{r}

Price_n_Location_geo <- st_as_sf(
  Price_n_Location |> drop_na(LATITUDE),
  coords = c('LONGITUDE', 'LATITUDE'),
  crs = st_crs(CensusTracts)
)

Price_n_Location_geo

```


```{r}

AffordableHousing_geo <- st_as_sf(
  AffordableHousing |> drop_na(LATITUDE),
  coords = c('LONGITUDE', 'LATITUDE'),
  crs = st_crs(CensusTracts)
)

AffordableHousing_geo

```

Side note: Initially constructed in 1974. After rennovation, it reopened (placed in service) in 2013. 
https://www.yardimatrix.com/property-types/multifamily/nashville/the-park-at-richards-road-850-richards-road-tn-37013--79489




```{r}
Price_n_Location <- Price_n_Location |>  
                      add_column(Distance_to_AffordableHousingComp = st_distance(Price_n_Location_geo, AffordableHousing_geo))
Price_n_Location
```

```{r}
Price_n_Location <- Price_n_Location |> 
                      rename(Sale_Price = price, Sale_Date = ownerdate) # |> 
                     # mutate(Sale_Year = year(ymd(Sale_Date)))
Price_n_Location
```


```{r}
HousesWithin2000ft <- Price_n_Location |> 
                      mutate_at(c('Distance_to_AffordableHousingComp'), as.numeric) |> 
                      filter(Distance_to_AffordableHousingComp <= 609.6) |> 
                      mutate(Within2000ft = 1)
HousesWithin2000ft
```


```{r}
HousesWithin2000ft_5to1YrBeforeOpening <- HousesWithin2000ft |> 
                                            filter(Sale_Date >= '2007-01-01') |> 
                                            filter(Sale_Date < '2012-01-01') |> 
                                            mutate(AfterOpening_1to5Yr = 0)
HousesWithin2000ft_5to1YrBeforeOpening
```

```{r}
HousesWithin2000ft_1to5YrAfterOpening <- HousesWithin2000ft |> 
                                            filter(Sale_Date > '2014-12-31') |> 
                                            filter(Sale_Date <= '2019-12-31') |>  
                                            mutate(AfterOpening_1to5Yr = 1)
HousesWithin2000ft_1to5YrAfterOpening
```



```{r}
HousesBetween2000n3000ft <- Price_n_Location |> 
                      mutate_at(c('Distance_to_AffordableHousingComp'), as.numeric) |> 
                      filter(Distance_to_AffordableHousingComp <= 914.4) |> 
                      filter(Distance_to_AffordableHousingComp >= 609.6) |> 
                      mutate(Within2000ft = 0)
HousesBetween2000n3000ft
```
```{r}
HousesBetween2000n3000ft_5to1YrBeforeOpening <- HousesBetween2000n3000ft |> 
                                            filter(Sale_Date >= '2007-01-01') |> 
                                            filter(Sale_Date < '2012-01-01') |> 
                                            mutate(AfterOpening_1to5Yr = 0)
HousesBetween2000n3000ft_5to1YrBeforeOpening
```

```{r}
HousesBetween2000n3000ft_1to5YrAfterOpening <- HousesBetween2000n3000ft |> 
                                            filter(Sale_Date > '2014-12-31') |> 
                                            filter(Sale_Date <= '2019-12-31') |> 
                                            mutate(AfterOpening_1to5Yr = 1)
HousesBetween2000n3000ft_1to5YrAfterOpening
```






```{r}
BinaryVarsAssignedTibble <- bind_rows(HousesWithin2000ft_5to1YrBeforeOpening, HousesWithin2000ft_1to5YrAfterOpening, HousesBetween2000n3000ft_5to1YrBeforeOpening, HousesBetween2000n3000ft_1to5YrAfterOpening)

BinaryVarsAssignedTibble
```


Difference-in-Difference analysis of prices of houses within 2000ft of affordable housing complex TNA20130015 vs. prices of houses between 2000ft and 3000ft from affordable housing complex TNA20130015 after its grand reopening did not yield statistically significant result. 
 
```{r}

lr_HousingPrice_DiffinDiff_nearAffordableApt <- glm(Sale_Price ~ Within2000ft + AfterOpening_1to5Yr + Within2000ft:AfterOpening_1to5Yr, family = gaussian(link = "identity"), data = BinaryVarsAssignedTibble)

summary(lr_HousingPrice_DiffinDiff_nearAffordableApt)

```
```{r}

lr_HousingPrice_DiffinDiff_nearAffordableApt <- glm(Sale_Price ~ Within2000ft + I(year(Sale_Date) - 2013) + Within2000ft:I(year(Sale_Date) - 2013) , family = gaussian(link = "identity"), data = BinaryVarsAssignedTibble)

summary(lr_HousingPrice_DiffinDiff_nearAffordableApt)

```




```{r}

lr_HousingPrice_DiffinDiff_nearAffordableApt <- glm(Sale_Price ~ Within2000ft + I(year(Sale_Date) - 2013) + Within2000ft:I(year(Sale_Date) - 2013):AfterOpening_1to5Yr, family = gaussian(link = "identity"), data = BinaryVarsAssignedTibble)

summary(lr_HousingPrice_DiffinDiff_nearAffordableApt)

```


```{r}
HousesWithin2000ft_1995to1YrBeforeOpening <- HousesWithin2000ft |> 
                                        #    filter(Sale_Date >= '2007-01-01') |> 
                                            filter(Sale_Date < '2012-01-01') |> 
                                            mutate(AfterOpening_1YrPlus = 0)
HousesWithin2000ft_1995to1YrBeforeOpening
```


```{r}
HousesWithin2000ft_1YrPlusAfterOpening <- HousesWithin2000ft |> 
                                            filter(Sale_Date > '2014-12-31') |> 
                                        #    filter(Sale_Date <= '2019-12-31') |>  
                                            mutate(AfterOpening_1YrPlus = 1)
HousesWithin2000ft_1YrPlusAfterOpening
```




```{r}
HousesBetween2000n3000ft_1995to1YrBeforeOpening <- HousesBetween2000n3000ft |> 
                                      #      filter(Sale_Date >= '2007-01-01') |> 
                                            filter(Sale_Date < '2012-01-01') |> 
                                            mutate(AfterOpening_1YrPlus = 0)
HousesBetween2000n3000ft_1995to1YrBeforeOpening
```




```{r}
HousesBetween2000n3000ft_1YrPlusAfterOpening <- HousesBetween2000n3000ft |> 
                                            filter(Sale_Date > '2014-12-31') |> 
                                 #           filter(Sale_Date <= '2019-12-31') |> 
                                            mutate(AfterOpening_1YrPlus = 1)
HousesBetween2000n3000ft_1YrPlusAfterOpening
```


```{r}
BinaryVarsAssignedTibble <- bind_rows(HousesWithin2000ft_1995to1YrBeforeOpening, HousesWithin2000ft_1YrPlusAfterOpening, HousesBetween2000n3000ft_1995to1YrBeforeOpening, HousesBetween2000n3000ft_1YrPlusAfterOpening)  

BinaryVarsAssignedTibble
```


"Did the reopening of affordable apartments affect nearby property values differently than properties further away, controlling for general time trends?"


RDD-Diff-in-Diff Hybrid running on the chassis of Diff-in-Diff:
In a very targeted way, we specifically care about prices of houses 2000ft within Affordable Housing Complex/Apt after the Reopening of Affordable Housing Complex/Apt. 

```{r}

# lr_HousingPrice_RDD_DiffinDiff_nearAffordableApt <- glm(Sale_Price ~ Within2000ft + I(Sale_Year - 2013) + Within2000ft:I(Sale_Year - 2013):AfterOpening_1YrPlus, family = gaussian(link = "identity"), data = BinaryVarsAssignedTibble)
# 
# summary(lr_HousingPrice_RDD_DiffinDiff_nearAffordableApt)

```


```{r}

lr_HousingPrice_RDD_DiffinDiff_nearAffordableApt <- glm(Sale_Price ~ Within2000ft + I(year(Sale_Date) - 2013) + Within2000ft:I(year(Sale_Date) - 2013):AfterOpening_1YrPlus, family = gaussian(link = "identity"), data = BinaryVarsAssignedTibble)

summary(lr_HousingPrice_RDD_DiffinDiff_nearAffordableApt)

```






Part 2 -

Given that the above affordable housing complex/apartment was a rebuild (instead of new construction), let's examine impact on pricing by a newly constructed affordable housing complex/apartment whose in service date precedes the 2008 real estate crash:

```{r}
NewlyConstructedAffordableHousing <- LIHTC |> filter(HUD_ID=='TNA20010150')
NewlyConstructedAffordableHousing 
```


```{r}
NewlyConstructedAffordableHousing_geo <- st_as_sf(
  NewlyConstructedAffordableHousing |> drop_na(LATITUDE),
  coords = c('LONGITUDE', 'LATITUDE'),
  crs = st_crs(CensusTracts)
)

NewlyConstructedAffordableHousing_geo
```

```{r}
Price_n_Location <- Price_n_Location |>  
                      add_column(Distance_to_TNA20010150 = st_distance(Price_n_Location_geo, NewlyConstructedAffordableHousing_geo))
Price_n_Location
```


```{r}
HousesWithin2000ft <- Price_n_Location |> 
                      mutate_at(c('Distance_to_TNA20010150'), as.numeric) |> 
                      filter(Distance_to_TNA20010150 <= 609.6) |> 
                      mutate(Within2000ft = 1)
HousesWithin2000ft
```


```{r}
HousesWithin2000ft_5to1YrBeforeOpening <- HousesWithin2000ft |> 
                                            filter(Sale_Date >= '1995-01-01') |> 
                                            filter(Sale_Date < '2000-01-01') |> 
                                            mutate(AfterOpening_1to5Yr = 0)
HousesWithin2000ft_5to1YrBeforeOpening
```

```{r}
HousesWithin2000ft_1to5YrAfterOpening <- HousesWithin2000ft |> 
                                            filter(Sale_Date > '2002-12-31') |> 
                                            filter(Sale_Date <= '2007-12-31') |> 
                                            mutate(AfterOpening_1to5Yr = 1)
HousesWithin2000ft_1to5YrAfterOpening
```





```{r}
HousesBetween2000n3000ft <- Price_n_Location |> 
                      mutate_at(c('Distance_to_TNA20010150'), as.numeric) |> 
                      filter(Distance_to_TNA20010150 <= 914.4) |> 
                      filter(Distance_to_TNA20010150 >= 609.6) |> 
                      mutate(Within2000ft = 0)
HousesBetween2000n3000ft
```

```{r}
HousesBetween2000n3000ft_5to1YrBeforeOpening <- HousesBetween2000n3000ft |> 
                                            filter(Sale_Date >= '2007-01-01') |> 
                                            filter(Sale_Date < '2012-01-01') |> 
                                            mutate(AfterOpening_1to5Yr = 0)
HousesBetween2000n3000ft_5to1YrBeforeOpening
```





```{r}
HousesBetween2000n3000ft_1to5YrAfterOpening <- HousesBetween2000n3000ft |> 
                                            filter(Sale_Date > '2002-12-31') |> 
                                            filter(Sale_Date <= '2007-12-31') |> 
                                            mutate(AfterOpening_1to5Yr = 1)
HousesBetween2000n3000ft_1to5YrAfterOpening
```


```{r}
BinaryVarsAssignedTibble_NewlyConstructed <- bind_rows(HousesWithin2000ft_5to1YrBeforeOpening, HousesWithin2000ft_1to5YrAfterOpening, HousesBetween2000n3000ft_5to1YrBeforeOpening, HousesBetween2000n3000ft_1to5YrAfterOpening)

BinaryVarsAssignedTibble_NewlyConstructed
```



Difference-in-Difference analysis of prices of houses within 2000ft of affordable housing complex TNA20010150 vs. prices of houses between 2000ft and 3000ft from affordable housing complex TNA20010150 after its opening yielded statistically significant positive effect - meaning that houses within 2000ft of the affordable housing complex TNA20010150 on averaged transacted at $66704 higher than prices of houses situated between 2000ft and 3000ft from affordable housing complex TNA20010150. 
 
```{r}

lr_HousingPrice_DiffinDiff_nearNewlyConstructedAffordableApt <- glm(Sale_Price ~ Within2000ft + AfterOpening_1to5Yr + Within2000ft:AfterOpening_1to5Yr, family = gaussian(link = "identity"), data = BinaryVarsAssignedTibble_NewlyConstructed)

summary(lr_HousingPrice_DiffinDiff_nearNewlyConstructedAffordableApt)

```


```{r}

lr_HousingPrice_DiffinDiff_nearNewlyConstructedAffordableApt <- glm(Sale_Price ~ Within2000ft + I(year(Sale_Date) - 2001) + Within2000ft:I(year(Sale_Date) - 2001), family = gaussian(link = "identity"), data = BinaryVarsAssignedTibble_NewlyConstructed)

summary(lr_HousingPrice_DiffinDiff_nearNewlyConstructedAffordableApt)

```











```{r}
HousesWithin2000ft_1995to1YrBeforeOpening <- HousesWithin2000ft |> 
                                        #filter(Sale_Date >= '1995-01-01') |> 
                                            filter(Sale_Date < '2000-01-01') |> 
                                            mutate(AfterOpening_1YrPlus = 0)
HousesWithin2000ft_1995to1YrBeforeOpening
```


```{r}
HousesWithin2000ft_1YrPlusAfterOpening <- HousesWithin2000ft |> 
                                            filter(Sale_Date > '2002-12-31') |> 
                                           # filter(Sale_Date <= '2007-12-31') |>  
                                            mutate(AfterOpening_1YrPlus = 1)
HousesWithin2000ft_1YrPlusAfterOpening
```




```{r}
HousesBetween2000n3000ft_1995to1YrBeforeOpening <- HousesBetween2000n3000ft |> 
                                      #      filter(Sale_Date >= '1995-01-01') |> 
                                            filter(Sale_Date < '2000-01-01') |> 
                                            mutate(AfterOpening_1YrPlus = 0)
HousesBetween2000n3000ft_1995to1YrBeforeOpening
```




```{r}
HousesBetween2000n3000ft_1YrPlusAfterOpening <- HousesBetween2000n3000ft |> 
                                            filter(Sale_Date > '2002-12-31') |> 
                                           # filter(Sale_Date <= '2007-12-31') |> 
                                            mutate(AfterOpening_1YrPlus = 1)
HousesBetween2000n3000ft_1YrPlusAfterOpening
```


```{r}
BinaryVarsAssignedTibble <- bind_rows(HousesWithin2000ft_1995to1YrBeforeOpening, HousesWithin2000ft_1YrPlusAfterOpening, HousesBetween2000n3000ft_1995to1YrBeforeOpening, HousesBetween2000n3000ft_1YrPlusAfterOpening)  

BinaryVarsAssignedTibble
```


"Did the reopening of affordable apartments affect nearby property values differently than properties further away, controlling for general time trends?"


RDD-Diff-in-Diff Hybrid running on the chassis of Diff-in-Diff:
In a very targeted way, we specifically care about prices of houses 2000ft within Affordable Housing Complex/Apt after the Reopening of Affordable Housing Complex/Apt. 

```{r}

lr_HousingPrice_RDD_nearNewlyConstructedAffordableApt <- glm(Sale_Price ~ Within2000ft + I(year(Sale_Date) - 2001) + Within2000ft:I(year(Sale_Date) - 2001):AfterOpening_1YrPlus, family = gaussian(link = "identity"), data = BinaryVarsAssignedTibble)

summary(lr_HousingPrice_RDD_nearNewlyConstructedAffordableApt)

```




